### 2020-04-24 ms_8.Rmd
joined data

---
title: "ms_8"
author: "Rachel Phan"
date: "4/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(readr)
library(readxl)
library(lubridate)
library(stringr)
library(gganimate)
library(ggplot2)
library(usmap)
library(maps)
library(gganimate)
library(scales)
```

```{r loading data, include=FALSE}

# Obesity data from https://chronicdata.cdc.gov/Nutrition-Physical-Activity-and-Obesity/Nutrition-Physical-Activity-and-Obesity-Behavioral/hn4x-zwk7/data (Center for Disease Control)

obesity_x <- read_csv("Nutrition__Physical_Activity__and_Obesity_-_Behavioral_Risk_Factor_Surveillance_System.csv")

# Food access data from https://www.ers.usda.gov/data-products/food-access-research-atlas/ (United States Department of Agriculture)

food_access_x <- read_excel("Food Accessibility.xlsx", sheet = 3)

# Rural income data from https://www.ers.usda.gov/data-products/atlas-of-rural-and-small-town-america/ (United States Department of Agriculture)

rural_income <- read_csv("Rural Atlas Data/Income.csv")

# Income data for all US Counties from https://www.ers.usda.gov/data-products/county-level-data-sets/download-data/ (US Bureau of Labour Statistics)

income <- read_excel("Unemployment.xls", skip = 4)

# Consumer Price Index ata from https://www.ers.usda.gov/data-products/food-price-outlook/ (United States Department of Agriculture)

cpi <- read_excel("historicalcpi.xlsx", skip = 1)

# What I want to observe here is how different variables affect obesity in the US - income, food accessibility, year, ethnicity, food price, which users can choose from when they pick the graph. Then I want to map out obesity/BMI onto a map of the US. I

# If obesity is affected by food access, then can look deeper into food access. How does ruralness affect food access? How do consumer food price indexes affect food accessibility? How does SNAP policy affect food access (can look at SNAP data from USDA: https://www.ers.usda.gov/data-products/snap-policy-data-sets/).

```

```{r clean obesity data, include=FALSE}

obesity_x <- obesity_x %>% 
    select(YearStart, LocationAbbr, Class, Data_Value, Low_Confidence_Limit, High_Confidence_Limit,
           Sample_Size, "Age(years)", Education, Gender, Income, "Race/Ethnicity")

obesity <- obesity_x %>% 
    rename("state" = LocationAbbr,
           "year" = YearStart,
           "perc_obese" = Data_Value,
           "conf.low" = Low_Confidence_Limit,
           "conf.high" = High_Confidence_Limit,
           "sample_size" = Sample_Size,
           "age" = "Age(years)",
           "education" = Education,
           "gender" = Gender,
           "class" = Class,
           "race" = "Race/Ethnicity",
           "income" = Income) %>% 
    filter(class == "Obesity / Weight Status",
           !is.na(perc_obese)) %>% 
    select(-class)

```

```{r obesity mean by year, include=FALSE}

# Maybe choose a state to look at its obesity rate change over time

obesity_mean <- obesity %>% 
    group_by(state, year) %>% 
    summarise(mean_perc = mean(perc_obese)) %>% 
    filter(state != "US")

obesity_mean_plot <- plot_usmap(data = obesity_mean, values = "mean_perc", color = "red", labels = FALSE) +
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5)) + 
    scale_fill_continuous(low = "white", high = "red", name = "Percent Obese") +
    labs(title = "Percent of Population (age 18+) Classified as Obese or Overweight in the US \n Year:
         {as.integer(frame_time)}",
         subtitle = "Percent of population tends to increase as time goes on",
         caption = "Data from Center for Disease Control") +
    theme_void() +
    transition_time(year)

animate(obesity_mean_plot)

anim_save("plots/obesity_mean_usmap.gif", plot = last_animation())

```

```{r obesity by region, include=FALSE}

obesity_region <- obesity_mean %>% 
    mutate(region = case_when(
        state %in% c("CT", "MA", "ME", "NH", "RI", "VT", "NJ", "NY", "PA") ~ "Northeast",
        state %in% c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD") ~ "Midwest",
        state %in% c("DC", "DE", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", 
                     "LA", "OK", "TX") ~ "South",
        state %in% c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA") ~ "West")
    ) %>% 
    mutate(year = as.numeric(year)) %>% 
  mutate(mean_perc = round(mean_perc, digits = 0))

obesity_region_plot <- obesity_region %>% 
    filter(!is.na(region)) %>%
    ggplot(aes(color = region, x = year, y = mean_perc)) +
    geom_jitter(aes(alpha = 0.1), show.legend = FALSE) +
    theme_minimal() +
    labs(title = "Percent of Population Classified as Obese or Overweight \n in US Over Time by Region",
         caption = "Data from Center for Disease Control",
         x = "Mean Percent of Population",
         y = "Year",
         color = "Region") +
    geom_smooth(method = "lm", se = FALSE, formula = y ~ x) +
    facet_wrap(~ region)

obesity_region_plot

ggsave("plots/obesity_region_plot.png", plot = last_plot())

```

```{r clean food access data, include=FALSE}

food_access_x <- food_access_x %>% 
    select(State, County, Urban, POP2010, LILATracts_1And10, PovertyRate, MedianFamilyIncome, 
           LA1and10, LAPOP1_10, LALOWI1_10)

food_access_y <- food_access_x %>% 
    rename("state" = State,
           "county" = County,
           "popul" = POP2010,
           "urban" = Urban,
           "poverty" = PovertyRate,
           "med_income" = MedianFamilyIncome,
           "lia_1_10" = LILATracts_1And10,
           "la_1_10" = LA1and10,
           "lia_pop"= LALOWI1_10,
           "la_pop" = LAPOP1_10) %>% 
    mutate(urban = as.factor(urban),
           lia_1_10 = as.factor(lia_1_10),
           la_1_10 = as.factor(la_1_10))

food_access_y$state = state.abb[match(food_access_y$state, state.name)]

food_access <- food_access_y %>% 
    mutate(region = case_when(
        state %in% c("CT", "MA", "ME", "NH", "RI", "VT", "NJ", "NY", "PA") ~ "Northeast",
        state %in% c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD") ~ "Midwest",
        state %in% c("DC", "DE", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", 
                     "LA", "OK", "TX") ~ "South",
        state %in% c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA") ~ "West")
    ) %>% 
    select(-county)

```

```{r mean poverty by state, include = FALSE}

food_poverty_mean <- food_access %>%
    filter(!is.na(region)) %>% 
    group_by(state) %>% 
    summarise(mean_poverty = mean(poverty))

food_poverty_mean_plot <- plot_usmap(data = food_poverty_mean, values = "mean_poverty", 
                                     color = "green", labels = FALSE) +
    theme(legend.position = "right",
          plot.title = element_text(hjust = 0.5)) + 
    scale_fill_continuous(low = "white", high = "green", name = "Mean Poverty Rate") +
    labs(title = "Poverty Rate in the US",
         caption = "Data from US Department of Agriculture and 2010 Census Bureau") +
    theme_void()
food_poverty_mean_plot

ggsave("plots/food_poverty_plot.png", plot = last_plot())

```

```{r food access lia distribution, include=FALSE}

food_lia <- food_access %>%
    filter(!is.na(region)) %>% 
    group_by(state, region) %>% 
    summarise(perc = sum(lia_1_10 == 1)/n()) %>% 
    mutate(perc = perc*100) %>% 
    mutate(perc = round(as.numeric(perc), digits = 0)) %>% 
    arrange(perc)

# Maybe could input the region in Shiny and people can play around with that?

food_lia_plot <- food_lia %>% 
    group_by(region) %>% 
    ggplot(aes(y = reorder(state, perc), x = perc, fill = region)) +
    geom_col() +
    theme_minimal() +
    theme(axis.text.y = element_text(hjust = 1, size = 5.5),
          panel.grid.major.y = element_line(color = "white"),
          panel.grid.minor.y = element_line(color = "white"),
          plot.subtitle = element_text(size = 8)) +
    labs(title = "Percent of Counties with a Lack of Food Access and Low Income Population",
         subtitle = "Lack of Food Access Defined by 33% of the Population Living More Than \n 1 Mile (Urban) Or 10 Miles (Rural) Away from Supermarket, Supercenter, or Grocery Store Who are Low Income",
         caption = "Data from US Department of Agriculture and 2010 Census Bureau",
         x = "Percent of Counties",
         y = "State",
         fill = "Region") +
    geom_text(aes(label = perc), hjust = -0.3, size = 2)

food_lia_plot

ggsave("plots/food_lia_plot.png", plot = last_plot())

```

```{r join obesity and food access, include=FALSE}

# compare % of people that're obese to % of people that're lia and la

obesity <- obesity %>% 
     mutate(region = case_when(
        state %in% c("CT", "MA", "ME", "NH", "RI", "VT", "NJ", "NY", "PA") ~ "Northeast",
        state %in% c("IL", "IN", "MI", "OH", "WI", "IA", "KS", "MN", "MO", "NE", "ND", "SD") ~ "Midwest",
        state %in% c("DC", "DE", "FL", "GA", "MD", "NC", "SC", "VA", "WV", "AL", "KY", "MS", "TN", "AR", 
                     "LA", "OK", "TX") ~ "South",
        state %in% c("AZ", "CO", "ID", "MT", "NV", "NM", "UT", "WY", "AK", "CA", "HI", "OR", "WA") ~ "West")
    ) %>% 
    mutate(year = as.numeric(year))

obesity_2011 <- obesity %>% 
    filter(year == 2011) %>%
    group_by(state, age, gender, income, race) %>% 
    summarise(mean_perc_obese = round(mean(perc_obese), digits = 0)) %>% 
    filter(state != "US" & state != "DC")

food_sum <- food_access %>% 
    group_by(state, urban) %>% 
    mutate(total_pop = sum(popul),
           total_lia_pop = sum(lia_pop),
           total_la_pop = sum(la_pop),
           perc_lia = round((total_lia_pop/total_pop)*100, digits = 0),
           perc_la = round((total_la_pop/total_pop)*100, digits = 0)) %>% 
    group_by(state, perc_lia, perc_la, urban) %>% 
    summarise(mean_poverty = round(mean(poverty), digits = 0)) %>% 
    filter(!is.na(state))

food_sum <- food_access %>% 
    group_by(state, urban, lia_pop, la_pop, region) %>% 
    summarise(total_pop = sum(popul),
           total_lia_pop = sum(lia_pop),
           total_la_pop = sum(la_pop),
           perc_lia = round((total_lia_pop/total_pop)*100, digits = 0),
           perc_la = round((total_la_pop/total_pop)*100, digits = 0),
           mean_poverty = round(mean(poverty), digits = 0)) %>% 
    select(-total_pop, -total_lia_pop, -total_la_pop, -la_pop, -lia_pop)

joined <- left_join(obesity_2011, food_sum, by = "state")

```

Shiny App

 fluidPage(

    titlePanel("GOV1005 Final Project"),

    navlistPanel("Header A",
                 tabPanel("Component 1"),
                 "Header B",
                 tabPanel("Component 3"),
                 tabPanel("Component 4")
    ),

    sidebarLayout(position = "right",
        sidebarPanel(img(src = "cartoon.png", height = 140, width = 200)),
        mainPanel(h3("Red Meat Production in the US Over Time"),
                  imageOutput("meat_clean_plot"),
                  br(),
                  br(),
                  br(),
                  br(),
                  br(),
                  br(),
                  actionButton("action", "Click")
    )
)

)

# Define server logic required to draw a histogram
server <- function(input, output) {

    output$meat_clean_plot <- renderImage({
        list(src = "meat_clean_plot.png",
             contentType = 'image/png',
             width = 550,
             height = 500)},
        deleteFile = FALSE
    )  
}